---
title: "Tutorial"
author: "Anton Bossenbroek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette gives explains some basic functionalities of the _GeneralTree_
package.

# Basic Tree Operations
In this section we will explain how to perform basic operations on a tree.

## Creating a tree
A General Tree is a R6 object. As such it can be created with the `new` method.
You should pass to `new` the `id` and `data` of the root node. Note that there
is currently no requirement that the data in the tree should be the same and/or
unique.

```{r}
# Create a root node with id = 0 and data = "root"
require(GeneralTree)
tree <- GeneralTree$new(id = 0, data = "root")
```

we can print the tree at any time to inspect its content,
```{r}
print(tree)
```

## Adding child nodes
Once the tree is created we can add child nodes to the tree. We can either call
`addNode` where we will have to specify the `id` of the parent node or we can
use `search` to find the parent node and add a child directly.

```{r}
# Add a child by specifying the parent.
tree$addNode(parent = 0, id = 1, data = "child0.1")
print(tree)

# Add a child by searching its parent.
tree$searchNode(1)$addChild(id = 2, data = "child1.2")
print(tree)
```

## Adding siblings
Siblings are created automatically when you are adding a new child to a node
that already has a child or when you explicity call `addSibling`.

```{r}
# Add a sibling by specifying the parent.
tree$addNode(parent = 0, id = 3, data = "child0.3")
print(tree)

# Add a sibling by searching its parent.
tree$searchNode(1)$addSibling(id = 4, data = "child0.4")
print(tree)
```

## Searching
There are two functions that help you retrieve data stored in the tree. Each
method returns a different object,

 * `searchNode` searches for a node.
 * `searchData` searches for data in the tree.

```{r}
# Let us create a mixed tree.
tree <- GeneralTree$new(0, "parent1")
tree$addNode(0, "a", "child.a")
tree$addNode(0, "b", "child.b")
tree$addNode("b", "c", "child.b.c")
tree$addNode("b", "d", "child.b.d")
tree$addNode("c", "e", "child.c.e")
tree$addNode("c", "f", "child.c.e")
```

Search the node with id `f`,
```{r}
print(tree$searchNode("f"))
```

Search the data with id `e`,
```{r}
tree$searchData("e")
```
