---
title: "Tutorial"
author: "Anton Bossenbroek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette gives explains some basic functionalities of the _GeneralTree_
package.

# Basic Tree Operations
In this section we will explain how to perform basic operations on a tree.

## Creating a tree
A General Tree is a R6 object. As such it can be created with the `new` method.
You should pass to `new` the `id` and `data` of the root node. Note that there
is currently no requirement that the data in the tree should be the same and/or
unique.

```{r}
# Create a root node with id = 0 and data = "root"
require(GeneralTree)
tree <- GeneralTree$new(id = 0, data = "root")
```

we can print the tree at any time to inspect its content,
```{r}
print(tree)
```

## Adding child nodes
Once the tree is created we can add child nodes to the tree. We can either call
`addNode` where we will have to specify the `id` of the parent node or we can
use `search` to find the parent node and add a child directly.

```{r}
# Add a child by specifying the parent.
tree$addNode(parent = 0, id = 1, data = "child0.1")
print(tree)

# Add a child by searching its parent.
tree$searchNode(1)$addChild(id = 2, data = "child1.2")
print(tree)
```

## Adding siblings
Siblings are created automatically when you are adding a new child to a node
that already has a child or when you explicity call `addSibling`.

```{r}
# Add a sibling by specifying the parent.
tree$addNode(parent = 0, id = 3, data = "child0.3")
print(tree)

# Add a sibling by searching its parent.
tree$searchNode(1)$addSibling(id = 4, data = "child0.4")
print(tree)
```

## Searching
There are two functions that help you retrieve data stored in the tree. Each
method returns a different object,

 * `searchNode` searches for a node.
 * `searchData` searches for data in the tree.

```{r}
# Let us create a mixed tree.
tree <- GeneralTree$new(0, "parent1")
tree$addNode(0, "a", "child.a")
tree$addNode(0, "b", "child.b")
tree$addNode("b", "c", "child.b.c")
tree$addNode("b", "d", "child.b.d")
tree$addNode("c", "e", "child.c.e")
tree$addNode("c", "f", "child.c.f")
tree$addNode("b", "g", "child.b.g")
tree$addNode("b", "h", "child.b.h")
tree$addNode("c", 1, "child.c.1")
```

Search the node with id `f`,
```{r}
print(tree$searchNode("f"))
```

Search the data with id `e`,
```{r}
tree$searchData("e")
```

## Plotting a tree
If you want a graphically representation of a tree you can do so easily,

```{r}
plot(tree)
```

We can change what is plotted as well as the shapes of the diagram,

```{r}
plot(tree, what = "data", color = "coral1", shape = "oval")
```

# Advanced Topics

## Casting
In the current version you can cast a tree to a data frame and vice versa as
well as a parsed object.

### Data frame
We can use the generic casting functions to cast to a data frame,

```{r}
as.data.frame(tree)
```

the other way around is also easy,

```{r}
# Let us define a data frame,
test_tree_df <- data.frame(
    ID = c("root", "child1", "child2", "child3"),
    DATA = c("parent1", "data3.1", "data1.2", "data1.3"),
    PARENT = c(NA, "child3", "root", "root"), stringsAsFactors = FALSE)
test_tree_df
```

By default `as.GeneralTree` will search for columns `id`, `data` and `parent`
where parent should have a `NA` to identify the root node. In our case we have
different column names. This can be passed to `as.GeneralTree` as we see below,

```{r}
as.GeneralTree(test_tree_df, id = "ID", data = "DATA", parent = "PARENT")
```

### Parsed code
To inspect how R code is parsed you can cast parsed code to a GeneralTree. Note
GeneralTree will automatically create a dummy root to compensate for the fact
that sometimes multiple expressions are at root,

In this example all entries exist at the root,
```{r}
p <- parse(text = "tree <- GeneralTree$new(1, \"parent1\")
                   tree$addNode(1, 2, \"child.1.2\")
                   tree$addNode(2, 3, \"child.2.3\")",
           keep.source = TRUE)
print(as.GeneralTree(p), what = "data")
```

In this example all entries hang below `test_that`,
```{r}
p <- parse(text =
           "test_that(\"test that the tree_walker with while loop\", {
              tree <- GeneralTree$new(1, \"parent1\")
              tree$addNode(1, 2, \"child.1.2\")
              tree$addNode(2, 3, \"child.2.3\")
             })",
           keep.source = TRUE)
print(as.GeneralTree(p), what = "data")
```

## Iterating through the tree
There are two approaches to iterate through the tree. The first is built in
whereas the second relies on `foreach` and `iterators` packages.

### Internal
You can iterate through the tree by first inializing an iterator and then
calling next element on that iterator.

```{r}
# Let us inspect the tree first,
print(tree, what = "data")

# Make a backup of the tree,
old_tree <- tree

i <- tree$iterator()
while (!is.null(i)) {
    i$setData(paste("id:", i$id, "-data", i$data))
    i <- tryCatch(i$nextElem(), error = function(e) NULL)
}

print(tree, what = "data")
```


### Foreach
With foreach you can quickly extract the id and data of the tree in a depth
first fashion,

```{r}
require(iterators)
require(foreach)
itx <- iter(old_tree, by = "id")
ids_in_tree <- foreach(i = itx, .combine = c) %do% c(i)
ids_in_tree
```

The main benefit of using `iter` is that you can iterate through the tree in
parllel. If you only need to perform operations on the `id` or `data` of a node
we highly recommend to use the latter method to traverse the tree. See the
following benchmark,

```{r}
p <- parse(text = "
            tree <- GeneralTree$new(1, \"parent1\")
            tree$addNode(1, 2, \"child.1.2\")
            tree$addNode(2, 3, \"child.2.3\")
            tree$addNode(3, 4, \"child.3.4\")
            tree$addNode(3, 5, \"child.3.5\")
            tree$addNode(1, 7, \"child.1.7\")
            tree$addNode(1, 8, \"child.1.8\")
            tree$addNode(8, 9, \"child.8.9\")
            tree$addNode(9, 10, \"child.9.10\")
            tree$addNode(9, 11, \"child.9.11\")
            tree$addNode(9, 12, \"child.9.12\")
            tree$addNode(12, 13, \"child.12.13\")
            tree$addNode(8, 14, \"child.8.14\")
            tree$addNode(2, 6, \"child.2.6\")", keep.source = TRUE)
tree <- as.GeneralTree(p)

require(microbenchmark)

microbenchmark({
  i <- tree$iterator()
  ids_in_tree <- c()
  while (!is.null(i)) {
    ids_in_tree <- c(ids_in_tree, i$id)
    i <- tryCatch(i$nextElem(), error = function(e) NULL)
  }
})

require(foreach)
require(iterators)
require(doParallel)
nThreads <- detectCores(logical = TRUE)
cl <- makeCluster(nThreads)
registerDoParallel(cl)
microbenchmark({
  itx <- iter(tree, by = "id")
  ids_in_tree <- foreach(i = itx, .combine = c) %dopar% c(i)
})
```

Note that currently `foreach` iteration is slower than native iteration.
